Statistical Reporting Inconsistencies in Experimental Linguistics

### 2024-07-25 manuscript/manuscript.tex
added credit scores and CoI

\title{The (mis)reporting of statistical results in experimental
linguistics}

@@

\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0000-0000-0001}} 
\par{ }
\par{       }
\par{Correspondence concerning this article should be addressed to Timo
B. Roettger, Email: timo.roettger@gmail.com}
}

@@

Since data and statistical protocol sharing remains rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}), interested readers are left with trusting the authors that
the statistical analysis is run and reported correctly. However, the
three sets of indices in (1) have clearly defined mathematical
relationships and can thus be easily checked for consistency. An F test
with the specified degrees of freedom and a test-statistic of 3.88
should result in a p-value of 0.053 which is larger, not smaller, than
0.05. Possible reasons for this inconsistency are manifold: It could be
a typo of the comparison sign, i.e.~the authors meant to use \texttt{=}
or \texttt{\textgreater{}} rather than \texttt{\textless{}}. Any of the
numbers could contain a typo. Sometimes an error might indicate
erroneous rounding (e.g.~0.057 being rounded down to 0.05). Without
access to data and scripts, it remains unclear to the reader what has
caused this inconsistency. Such inconsistencies can be particularly
concerning if the calculated p-value (here 0.057) and the reported
p-values (here \textgreater0.05) are not on the same side of the alpha
threshold. In NHST, p-values below a conventionalized alpha threshold,
most commonly 0.05, are interpreted as evidence that the data are
sufficiently inconsistent with the null hypothesis (``significant'').
P-values above that threshold are considered consistent with the null
hypothesis and practically lead to rejecting the alternative hypothesis
(``non-significant''). In (1) above, the reported p-value suggest a
significant result, the p-value derived from the degrees of freedom and
the test statistic suggest a non-significant result. In the following we
refer to these inconsistencies as ``decision inconsistencies''.

@@

We used the R package wtatcheck (Version 1.4.1-beta.2,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using ``regular expressions''. That way, statcheck can
detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation
tests, and Q tests as long as the test result fulfills three conditions:
(a) the test result is reported completely including the test statistic,
the degrees of freedom (if applicable), and the p-value; (b) the test
result is in the body of the text, i.e.~Statcheck usually misses
reported in tables; and (c) the test result is reported in American
Psychological Association style (\citeproc{ref-APA2020}{American
Psychological Association, 2020}). Given these constraints, Statcheck is
estimated to detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-values and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

@@

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the term ``experiment* in title, abstract and/or
keywords. Out of these 100 journals, we selected all journals with at
least 10\% of articles containing the search string''experiment*``. Out
of the remaining 37 journals, we selected those journals that urged APA
formatting either in the main body of the text or specifically regarding
statistics in the author guidelines, resulting in nine remaining
journals. Moreover, to access the articles in .pdf format, the articles
had to be either accessible to us through our library license, or open
access, resulting in a final list of eight journals: Applied
Psycholinguistics (APS), Bilingualism: Language and Cognition (BLC),
Linguistic Approaches to Bilingualism (LAB), Language and Speech (LaS),
Language Learning and Techology (LLT), Journal of Language and Social
Psychology (LSP), Journal of Child Language (JCL), and Studies in Second
Language Acquisition (SLA).

We included only original research articles within the publication years
of 2000-2023, excluding any book reviews, response articles,
commentaries, editorials, corrigenda, errata, advertisements, etc.
Articles from LAB spanned 2011-2023, while the other journals spanned
2000-2023. Statcheck could not parse 157 articles, likely related to
issues with rendering the Chi-Squared symbol being erroneously converted
from .pdf to .txt. This procedure resulted in 5804 research articles
which were submitted to analysis.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0755}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1698}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1887}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1792}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1509}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2358}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Journal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
eligible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessible results
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
inconsistencies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
decision inconsistencies
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
APS & 953 & 690 & 9570 & 1368 & 170 \\
BLC & 964 & 610 & 9093 & 1161 & 120 \\
JCL & 1109 & 529 & 6240 & 750 & 69 \\
LAB & 471 & 133 & 1719 & 234 & 25 \\
LLT & 421 & 111 & 919 & 201 & 61 \\
LSP & 695 & 376 & 4320 & 429 & 60 \\
LaS & 598 & 363 & 4717 & 552 & 86 \\
SLA & 593 & 247 & 2954 & 471 & 64 \\
Total & 5804 & 3059 & 39532 & 5166 & 655 \\
\end{longtable}

@@

The proportion of inconsistencies ranged from 10 to 22\% across journals
(1 to 7\% for decisions inconsistencies) (see
Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
year of publication (see Figure~\ref{fig-year-prop}). On average, 50\%
of assessable articles contained one or more inconsistencies (journals
range from 43 to 56\%) and 13\% contained one or more decisions
inconsistencies (journals range from 11 to 15\%).

@@

When examining reported against recalculated p-values for all
inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
spatial patterns (only inconsistencies that are based on `=' comparisons
were included to make them easier to interpret, n = 3779 out of 5166).
First, the center of density of points is located in the bottom left
corner with more p-values reported closer to the standard alpha level of
0.05. This is not surprising, since the quantitative sciences exhibit a
documented publication bias, with hypotheses being more often confirmed
(and thus p \textless{} .05) than not
(\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
is a linear density of points along the diagonal axis which corresponds
to numerically small inconsistencies, some of which might be related to
simple rounding errors. However, comparing the diagonal to the black
line, which represents a linear model predicting recalculated by
reported p-values, we can see a clear divergence of what is expected if
inconsistencies were equally likely in both directions
(i.e.~recalculated p-values were as likely to be smaller than the
reported p-value as larger). The slope of the regression line is flatter
than the diagonal axis which means that, on average, inconsistencies
have a tendency to exhibit smaller reported p-values than their
recalculated counterparts. Similarly, inconsistencies that report the
p-value as being larger or smaller than a reference value (e.g.~p
\textgreater{} 0.05 or p \textless{} 0.05 ) are not equally prevalent.
There were 4.4\% of inconsistencies with p being reported as larger than
a reference but 7.4\% inconsistencies with p being reported as smaller
than a reference (e.g.~p \textless{} 0.05). So even if we assumed these
inconsistencies were merely typos of the comparison sign,
inconsistencies that erroneously report the p-value to be smaller than a
reference value are more frequent than inconsistencies that erroneously
report the p-value to be larger than a reference value. These biases are
also reflected in decisions inconsistencies. Of all decisions
inconsistencies (n = 655), 71\% represent cases in which a reported
significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05).

@@

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent such assessment using Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Previous
studies report on inconsistency rates between 4\% and 14\%, with between
10\% and 63\% of articles containing at least one inconsistency and
between 3\% and 21\% contain at least one decision consistencies.

Even if the prevalence of these inconsistencies could be largely
attributed to inconsequential typos or rounding errors (an assumption we
cannot test without access to the data), the sheer amount of these
inconsistencies that have made it through peer-review should concern us.
They are human errors. If such a substantial amount of errors is found
in plain site, the question naturally arises as to how many errors
during the data analysis itself remain undetected. If the tip of the
iceberg above water is already so large, how large is the body of the
iceberg underneath the surface?

@@

\bibitem[\citeproctext]{ref-arvan2022reproducibility}
Arvan, M., Pina, L., \& Parde, N. (2022). Reproducibility in
computational linguistics: Is source code enough? \emph{Proceedings of
the 2022 Conference on Empirical Methods in Natural Language
Processing}, 2350--2361.

\bibitem[\citeproctext]{ref-bakker2011mis}
Bakker, M., \& Wicherts, J. M. (2011). The (mis) reporting of
statistical results in psychology journals. \emph{Behavior Research
Methods}, \emph{43}, 666--678.

\bibitem[\citeproctext]{ref-bakker2014outlier}
Bakker, M., \& Wicherts, J. M. (2014). Outlier removal and the relation
with reporting errors and quality of psychological research. \emph{PloS
One}, \emph{9}(7), e103360.

\bibitem[\citeproctext]{ref-barber2017meticulous}
Barber, L. K. (2017). Meticulous manuscripts, messy results: Working
together for robust science reporting. \emph{Stress \& Health},
\emph{33}(2), 89--91.

\bibitem[\citeproctext]{ref-berle2007inconsistencies}
Berle, D., \& Starcevic, V. (2007). Inconsistencies between reported
test statistics and p-values in two psychiatry journals.
\emph{International Journal of Methods in Psychiatric Research},
\emph{16}(4), 202--207.

\bibitem[\citeproctext]{ref-bochynska2023reproducible}
Bochynska, A., Keeble, L., Halfacre, C., Casillas, J. V., Champagne,
I.-A., Chen, K., Röthlisberger, M., Buchanan, E. M., \& Roettger, T.
(2023). Reproducible research practices and transparency across
linguistics. \emph{Glossa Psycholinguistics}, \emph{2}(1).

\bibitem[\citeproctext]{ref-boland2017ten}
Boland, M. R., Karczewski, K. J., \& Tatonetti, N. P. (2017). Ten simple
rules to enable multi-site collaborations through data sharing. In
\emph{PLoS computational biology} (1; Vol. 13, p. e1005278). Public
Library of Science San Francisco, CA USA.

\bibitem[\citeproctext]{ref-buckley2023estimating}
Buckley, J., Hyland, T., \& Seery, N. (2023). Estimating the
replicability of technology education research. \emph{International
Journal of Technology and Design Education}, \emph{33}(4), 1243--1264.

\bibitem[\citeproctext]{ref-caperos2013consistency}
Caperos, J. M., \& Pardo, A. (2013). Consistency errors in p-values
reported in spanish psychology journals. \emph{Psicothema},
\emph{25}(3), 408--414.

\bibitem[\citeproctext]{ref-casillas2023opening}
Casillas, J. V., Constantin-Dureci, G., Rascón, I. A., Shao, J.,
Rodrı́guez, S. A., Gadamsetty, A., Minetti, A., Laungani, K., Thatcher,
J., Gardere, R.-T., et al. (2023). \emph{Opening open science to all:
Demystifying reproducibility and transparency practices in linguistic
research}.

\bibitem[\citeproctext]{ref-claesen2023data}
Claesen, A., Vanpaemel, W., Maerten, A.-S., Verliefde, T., Tuerlinckx,
F., \& Heyman, T. (2023). Data sharing upon request and statistical
consistency errors in psychology: A replication of wicherts, bakker and
molenaar (2011). \emph{Plos One}, \emph{18}(4), e0284243.

\bibitem[\citeproctext]{ref-colavizza2020citation}
Colavizza, G., Hrynaszkiewicz, I., Staden, I., Whitaker, K., \&
McGillivray, B. (2020). The citation advantage of linking publications
to research data. \emph{PloS One}, \emph{15}(4), e0230416.

\bibitem[\citeproctext]{ref-colombo2018statistical}
Colombo, M., Duev, G., Nuijten, M. B., \& Sprenger, J. (2018).
Statistical reporting inconsistencies in experimental philosophy.
\emph{PloS One}, \emph{13}(4), e0194360.

@@

\bibitem[\citeproctext]{ref-franco2014publication}
Franco, A., Malhotra, N., \& Simonovits, G. (2014). Publication bias in
the social sciences: Unlocking the file drawer. \emph{Science},
\emph{345}(6203), 1502--1505.

\bibitem[\citeproctext]{ref-fugelsang2004theory}
Fugelsang, J. A., Stein, C. B., Green, A. E., \& Dunbar, K. N. (2004).
Theory and data interactions of the scientific mind: Evidence from the
molecular and the cognitive laboratory. \emph{Canadian Journal of
Experimental Psychology/Revue Canadienne de Psychologie
Exp{é}rimentale}, \emph{58}(2), 86.

\bibitem[\citeproctext]{ref-garcia2004incongruence}
Garcı́a-Berthou, E., \& Alcaraz, C. (2004). Incongruence between test
statistics and p values in medical papers. \emph{BMC Medical Research
Methodology}, \emph{4}, 1--5.

\bibitem[\citeproctext]{ref-green2018statcheck}
Green, C. D., Abbas, S., Belliveau, A., Beribisky, N., Davidson, I. J.,
DiGiovanni, J., Heidari, C., Martin, S. M., Oosenbrug, E., \&
Wainewright, L. M. (2018). Statcheck in canada: What proportion of CPA
journal articles contain errors in the reporting of p-values?
\emph{Canadian Psychology/Psychologie Canadienne}, \emph{59}(3), 203.

\bibitem[\citeproctext]{ref-gross2021fidelity}
Groß, T. (2021). Fidelity of statistical reporting in 10 years of cyber
security user studies. \emph{Socio-Technical Aspects in Security and
Trust: 9th International Workshop, STAST 2019, Luxembourg City,
Luxembourg, September 26, 2019, Revised Selected Papers 9}, 3--26.

\bibitem[\citeproctext]{ref-hardwicke2018data}
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G.
C., Kidwell, M. C., Hofelich Mohr, A., Clayton, E., Yoon, E. J., Henry
Tessler, M., et al. (2018). Data availability, reusability, and analytic
reproducibility: Evaluating the impact of a mandatory open data policy
at the journal cognition. \emph{Royal Society Open Science},
\emph{5}(8), 180448.

>>>>

\title{Statistical Reporting Inconsistencies in Experimental
Linguistics}

@@

\authornote{\par{\addORCIDlink{Timo B. Roettger}{0000-0000-0000-0001}} 
\par{ }
\par{   The authors have no conflict of interest to
declare.   Conceptionalization, Methodology, Validation, Formal
Analysis, Review \& Editing of Manuscript, Data Curation - T.B.R. \&
D.L.J.E.; Software, Investigation - D.L.J.E.; Writing of Original Draft,
Visualization, Supervision - T.B.R. }
\par{Correspondence concerning this article should be addressed to Timo
B. Roettger, Email: timo.roettger@gmail.com}
}

@@

Since data and statistical protocol sharing remains rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}), interested readers are left with trusting the authors that
the statistical analysis is run and reported correctly. However, the
three sets of indices in (1) have clearly defined mathematical
relationships and can thus be easily checked for consistency. An F test
with the specified degrees of freedom and a test-statistic of 3.88
should result in a p-value of 0.053 which is larger, not smaller, than
0.05. Possible reasons for this inconsistency are manifold: It could be
a typo of the comparison sign, i.e.~the authors meant to use \texttt{=}
or \texttt{\textgreater{}} rather than \texttt{\textless{}}. Any of the
numbers could contain a typo. Sometimes an error might indicate
erroneous rounding (e.g.~0.057 being rounded down to 0.05). Without
access to data and scripts, it remains unclear to the reader what has
caused this inconsistency. Such inconsistencies can be particularly
concerning if the calculated p-value (here 0.057) and the reported
p-values (here \textless0.05) are not on the same side of the alpha
threshold. In NHST, p-values below a conventionalized alpha threshold,
most commonly 0.05, are interpreted as evidence that the data are
sufficiently inconsistent with the null hypothesis (``significant'').
P-values above that threshold are considered consistent with the null
hypothesis and practically lead to rejecting the alternative hypothesis
(``non-significant''). In (1) above, the reported p-value suggest a
significant result, the p-value derived from the degrees of freedom and
the test statistic suggest a non-significant result. In the following we
refer to these inconsistencies as ``decision inconsistencies''.

@@

We used the R package Statcheck (Version 1.4.1-beta.2,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2023}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using ``regular expressions''. That way, statcheck can
detect results of t-tests, F-tests, Z-tests, χ2-tests, correlation
tests, and Q tests as long as the test result fulfills three conditions:
(a) the test result is reported completely including the test statistic,
the degrees of freedom (if applicable), and the p-value; (b) the test
result is in the body of the text, i.e.~Statcheck usually misses
reported in tables; and (c) the test result is reported in American
Psychological Association style (\citeproc{ref-APA2020}{American
Psychological Association, 2020}). Given these constraints, Statcheck is
estimated to detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-values and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

@@

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the search string ``experiment*'' in title, abstract
and/or keywords. Out of these 100 journals, we selected all journals
with at least 10\% of articles containing the search string
``experiment*''. Out of the remaining 37 journals, we selected those
journals that urged APA formatting either in the main body of the text
or specifically regarding statistics in the author guidelines, resulting
in nine remaining journals. Moreover, to access the articles in .pdf
format, the articles had to be either accessible to us through our
library license, or open access, resulting in a final list of eight
journals: Applied Psycholinguistics (APS), Bilingualism: Language and
Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
and Speech (LaS), Language Learning and Techology (LLT), Journal of
Language and Social Psychology (LSP), Journal of Child Language (JCL),
and Studies in Second Language Acquisition (SLA).

We included only original research articles within the publication years
of 2000-2023, excluding any book reviews, response articles,
commentaries, editorials, corrigenda, errata, advertisements, etc.
Articles from LAB spanned 2011-2023, while the other journals spanned
2000-2023. Statcheck could not parse 157 articles, likely related to
issues with rendering the Chi-Squared symbol being erroneously converted
from .pdf to .txt. No assessable statistics were found by Statcheck in
JCL articles from 2000 to 2002 and in SLA articles from 2004 to 2008.
This procedure resulted in 5804 research articles which were submitted
to analysis.

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0755}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1698}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1887}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1792}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1509}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2358}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Journal
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
eligible articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable articles
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
assessable results
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
inconsistencies
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
decision inconsistencies
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
APS & 953 & 690 & 9570 & 1368 & 170 \\
BLC & 964 & 610 & 9093 & 1161 & 120 \\
JCL & 1109 & 529 & 6240 & 750 & 69 \\
LAB & 471 & 133 & 1719 & 234 & 25 \\
LLT & 421 & 111 & 919 & 201 & 61 \\
LSP & 695 & 376 & 4320 & 429 & 60 \\
LaS & 598 & 363 & 4717 & 552 & 86 \\
SLA & 593 & 247 & 2954 & 471 & 64 \\
Total & 5804 & 3059 & 39532 & 5166 & 655 \\
\end{longtable}

@@

The proportion of inconsistencies ranged from 10 to 22\% across journals
(1 to 7\% for decision inconsistencies) (see
Figure~\ref{fig-stacked-bar}). These rates appear to be stable across
year of publication (see Figure~\ref{fig-year-prop}). On average, 50\%
of assessable articles contained one or more inconsistencies (journals
range from 43 to 56\%) and 13\% contained one or more decision
inconsistencies (journals range from 11 to 15\%).

@@

When examining reported against recalculated p-values for all
inconsistencies (see Figure~\ref{fig-scatter}), we can identify certain
spatial patterns (only inconsistencies that are based on `=' comparisons
were included to make them easier to interpret, n = 3779 out of 5166).
First, the center of density of points is located in the bottom left
corner with more p-values reported closer to the standard alpha level of
0.05. This is not surprising, since the quantitative sciences exhibit a
documented publication bias, with hypotheses being more often confirmed
(and thus p \textless{} .05) than not
(\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}). Second, there
is a linear density of points along the diagonal axis which corresponds
to numerically small inconsistencies, some of which might be related to
simple rounding errors. However, comparing the diagonal to the black
line, which represents a linear model predicting recalculated by
reported p-values, we can see a clear divergence of what is expected if
inconsistencies were equally likely in both directions. The slope of the
regression line is flatter than the diagonal axis which means that, on
average, inconsistencies have a tendency to exhibit smaller reported
p-values than their recalculated counterparts. Similarly,
inconsistencies that report the p-value as being larger or smaller than
a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05 )
are not equally prevalent. There were 4.4\% of inconsistencies with p
being reported as larger than a reference but 7.4\% inconsistencies with
p being reported as smaller than a reference (e.g.~p \textless{} 0.05).
So even if we assumed these inconsistencies were merely typos of the
comparison sign, inconsistencies that erroneously report the p-value to
be smaller than a reference value are more frequent than inconsistencies
that erroneously report the p-value to be larger than a reference value.
These biases are also reflected in decision inconsistencies. Of all
decision inconsistencies (n = 655), 71\% represent cases in which a
reported significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05).

@@

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent such assessment using Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Previous
studies report on inconsistency rates between 4\% and 14\%, with between
10\% and 63\% of articles containing at least one inconsistency and
between 3\% and 21\% contain at least one decision consistency.

Even if the prevalence of these inconsistencies could be largely
attributed to inconsequential typos or rounding errors (an assumption we
cannot test without access to the data), the sheer amount of these
inconsistencies that have made it through peer-review should concern us.
They are human errors. If such a substantial amount of errors is found
in plain site, the question naturally arises as to how many errors
during the data analysis itself remain undetected. If the tip of the
iceberg is already so large, what is the volume of the submerged
iceberg?

@@

\bibitem[\citeproctext]{ref-arvan2022reproducibility}
Arvan, M., Pina, L., \& Parde, N. (2022). Reproducibility in
computational linguistics: Is source code enough? \emph{Proceedings of
the 2022 Conference on Empirical Methods in Natural Language
Processing}, 2350--2361.
\url{https://doi.org/10.18653/v1/2022.emnlp-main.150}

\bibitem[\citeproctext]{ref-bakker2011mis}
Bakker, M., \& Wicherts, J. M. (2011). The (mis) reporting of
statistical results in psychology journals. \emph{Behavior Research
Methods}, \emph{43}, 666--678.
\url{https://doi.org/10.3758/s13428-011-0089-5}

\bibitem[\citeproctext]{ref-bakker2014outlier}
Bakker, M., \& Wicherts, J. M. (2014). Outlier removal and the relation
with reporting errors and quality of psychological research. \emph{PloS
One}, \emph{9}(7), e103360.
\url{https://doi.org/10.1371/journal.pone.0103360}

\bibitem[\citeproctext]{ref-barber2017meticulous}
Barber, L. K. (2017). Meticulous manuscripts, messy results: Working
together for robust science reporting. \emph{Stress \& Health},
\emph{33}(2), 89--91. \url{https://doi.org/10.1002/smi.2756}

\bibitem[\citeproctext]{ref-berle2007inconsistencies}
Berle, D., \& Starcevic, V. (2007). Inconsistencies between reported
test statistics and p-values in two psychiatry journals.
\emph{International Journal of Methods in Psychiatric Research},
\emph{16}(4), 202--207. \url{https://doi.org/10.1002/mpr.225}

\bibitem[\citeproctext]{ref-bochynska2023reproducible}
Bochynska, A., Keeble, L., Halfacre, C., Casillas, J. V., Champagne,
I.-A., Chen, K., Röthlisberger, M., Buchanan, E. M., \& Roettger, T.
(2023). Reproducible research practices and transparency across
linguistics. \emph{Glossa Psycholinguistics}, \emph{2}(1).
\url{https://doi.org/10.5070/G6011239}

\bibitem[\citeproctext]{ref-boland2017ten}
Boland, M. R., Karczewski, K. J., \& Tatonetti, N. P. (2017). Ten simple
rules to enable multi-site collaborations through data sharing. In
\emph{PLoS computational biology} (1; Vol. 13, p. e1005278). Public
Library of Science San Francisco, CA USA.
\url{https://doi.org/10.1371/journal.pcbi.1005278}

\bibitem[\citeproctext]{ref-buckley2023estimating}
Buckley, J., Hyland, T., \& Seery, N. (2023). Estimating the
replicability of technology education research. \emph{International
Journal of Technology and Design Education}, \emph{33}(4), 1243--1264.
\url{https://doi.org/10.1007/s10798-022-09787-6}

\bibitem[\citeproctext]{ref-caperos2013consistency}
Caperos, J. M., \& Pardo, A. (2013). Consistency errors in p-values
reported in spanish psychology journals. \emph{Psicothema},
\emph{25}(3), 408--414. \url{https://doi.org/10.7334/psicothema2012.207}

\bibitem[\citeproctext]{ref-casillas2023opening}
Casillas, J. V., Constantin-Dureci, G., Rascón, I. A., Shao, J.,
Rodrı́guez, S. A., Gadamsetty, A., Minetti, A., Laungani, K., Thatcher,
J., Gardere, R.-T., et al. (2023). \emph{Opening open science to all:
Demystifying reproducibility and transparency practices in linguistic
research}. \url{https://doi.org/10.31234/osf.io/spz4w}

\bibitem[\citeproctext]{ref-claesen2023data}
Claesen, A., Vanpaemel, W., Maerten, A.-S., Verliefde, T., Tuerlinckx,
F., \& Heyman, T. (2023). Data sharing upon request and statistical
consistency errors in psychology: A replication of wicherts, bakker and
molenaar (2011). \emph{Plos One}, \emph{18}(4), e0284243.
\url{https://doi.org/10.1371/journal.pone.0284243}

\bibitem[\citeproctext]{ref-colavizza2020citation}
Colavizza, G., Hrynaszkiewicz, I., Staden, I., Whitaker, K., \&
McGillivray, B. (2020). The citation advantage of linking publications
to research data. \emph{PloS One}, \emph{15}(4), e0230416.
\url{https://doi.org/10.1371/journal.pone.0230416}

\bibitem[\citeproctext]{ref-colombo2018statistical}
Colombo, M., Duev, G., Nuijten, M. B., \& Sprenger, J. (2018).
Statistical reporting inconsistencies in experimental philosophy.
\emph{PloS One}, \emph{13}(4), e0194360.
\url{https://doi.org/10.1371/journal.pone.0194360}

@@

\bibitem[\citeproctext]{ref-franco2014publication}
Franco, A., Malhotra, N., \& Simonovits, G. (2014). Publication bias in
the social sciences: Unlocking the file drawer. \emph{Science},
\emph{345}(6203), 1502--1505.
\url{https://doi.org/10.1126/science.1255484}

\bibitem[\citeproctext]{ref-fugelsang2004theory}
Fugelsang, J. A., Stein, C. B., Green, A. E., \& Dunbar, K. N. (2004).
Theory and data interactions of the scientific mind: Evidence from the
molecular and the cognitive laboratory. \emph{Canadian Journal of
Experimental Psychology/Revue Canadienne de Psychologie
Exp{é}rimentale}, \emph{58}(2), 86.
\url{https://doi.org/10.1037/h0085799}

\bibitem[\citeproctext]{ref-garcia2004incongruence}
Garcı́a-Berthou, E., \& Alcaraz, C. (2004). Incongruence between test
statistics and p values in medical papers. \emph{BMC Medical Research
Methodology}, \emph{4}, 1--5.
\url{https://doi.org/10.1186/1471-2288-4-13}

\bibitem[\citeproctext]{ref-green2018statcheck}
Green, C. D., Abbas, S., Belliveau, A., Beribisky, N., Davidson, I. J.,
DiGiovanni, J., Heidari, C., Martin, S. M., Oosenbrug, E., \&
Wainewright, L. M. (2018). Statcheck in canada: What proportion of CPA
journal articles contain errors in the reporting of p-values?
\emph{Canadian Psychology/Psychologie Canadienne}, \emph{59}(3), 203.
\url{https://doi.org/10.1037/cap0000139}

\bibitem[\citeproctext]{ref-gross2021fidelity}
Groß, T. (2021). Fidelity of statistical reporting in 10 years of cyber
security user studies. \emph{Socio-Technical Aspects in Security and
Trust: 9th International Workshop, STAST 2019, Luxembourg City,
Luxembourg, September 26, 2019, Revised Selected Papers 9}, 3--26.
\url{https://doi.org/10.1007/978-3-030-55958-8_1}

\bibitem[\citeproctext]{ref-hardwicke2018data}
Hardwicke, T. E., Mathur, M. B., MacDonald, K., Nilsonne, G., Banks, G.
C., Kidwell, M. C., Hofelich Mohr, A., Clayton, E., Yoon, E. J., Henry
Tessler, M., et al. (2018). Data availability, reusability, and analytic
reproducibility: Evaluating the impact of a mandatory open data policy
at the journal cognition. \emph{Royal Society Open Science},
\emph{5}(8), 180448. \url{https://doi.org/10.1098/rsos.180448}
